Community Data Science meets Digital Heritage: Controversies about World Heritage Sites in the Online Community of Wikipedia

### 2019-05-06 code/paper.Rmd
stash

ip_regex <- "(?(?=.*?(\\d+\\.\\d+\\.\\d+\\.\\d+).*?)(\\1|))"

get_ip_addresses <- function(x){
  re <- regexpr(
  ip_regex, 
  x, perl = TRUE)
regmatches(x, re)
}

@@

Do WH sites have more geographically diverse contributions?

Get locations for random pages.

How many IP edits on each page?

```{r}
randoms_revision_history_page_details <- 
  map_df(ten_k_random_wp_pages_lst, 
         ~.x$revision_history_page_details, 
         .id = "page")

randoms_revision_history_page_details_ip <- 
randoms_revision_history_page_details %>% 
    mutate(is_ip_address = ifelse(get_ip_addresses(rh_user) != "", TRUE, FALSE),
         page = as.numeric(page))

randoms_revision_history_page_details_ip_prop <- 
  randoms_revision_history_page_details_ip %>% 
  group_by(page) %>% 
  summarise(prop_ip_addresses = sum(is_ip_address) / n() )

ggplot(randoms_revision_history_page_details_ip_prop,
       aes(prop_ip_addresses)) +
  geom_histogram() +
  theme_minimal()
```

Overall location of IP edits for all random pages

```{r}
randoms_revision_history_page_details_ip_location <- 
randoms_revision_history_page_details_ip %>% 
  filter(is_ip_address) %>% 
  mutate(ip_location = map(rh_user, ~maxmind(.x, ip_db)))  %>% 
  unnest(ip_location)

randoms_revision_history_page_details_ip_location_tally <- 
randoms_revision_history_page_details_ip_location %>% 
  filter(!is.na(country_name)) %>% 
  mutate(country_name = fct_rev(fct_infreq(country_name))) %>% 
  group_by(country_name) %>% 
  tally(sort = TRUE) %>% 
  filter(n > 100)

ggplot(randoms_revision_history_page_details_ip_location_tally,
       aes(country_name, n)) +
  geom_col() +
  coord_flip()   +
  theme_minimal() +
  scale_y_continuous(labels = scales::comma)
```

What's the geographic diversity of edits per random page?

```{r}
randoms_revision_history_page_details_ip_location_props_per_page <- 
randoms_revision_history_page_details_ip_location %>% 
  filter(!is.na(country_name)) %>% 
  mutate(country_name = fct_rev(fct_infreq(country_name))) %>% 
  group_by(page, country_name) %>% 
  tally() %>% 
  mutate(prop = n / sum(n)) %>% 
  filter(country_name %in% randoms_revision_history_page_details_ip_location_tally$country_name)

ggplot(randoms_revision_history_page_details_ip_location_props_per_page,
       aes(country_name, 
           prop)) +
  geom_boxplot() +
  coord_flip()

randoms_revision_history_page_details_ip_location_props_per_page_div <- 
  randoms_revision_history_page_details_ip_location_props_per_page %>% 
  group_by(page) %>% 
  summarise(shannon = vegan::diversity(n))

ggplot(randoms_revision_history_page_details_ip_location_props_per_page_div,
       aes( shannon)) +
  geom_histogram() 
```

```{r}
# analyse the countries of IP address vs country of WH site
revision_history_page_details_ip_addresses <- 
revision_history_page_details %>% 
  mutate(edits_from_ip_addresses = map(revision_history_page_details,
                          ~.x %>% 
                            filter(get_ip_addresses(rh_user) != "") %>% 
                            mutate(ip_location = map(rh_user, 
                                                     ~maxmind(.x, ip_db)))  %>% 
                            unnest(ip_location)))

revision_history_page_details_ip_addresses_df <- 
revision_history_page_details_ip_addresses %>% 
  unnest(edits_from_ip_addresses)

@@

@@

Look at distributions of variables, then say 2sd is extreme, and match those to the WH pages to find out what they are

What's the typical distribution of talk page length?

```{r}
ten_k_random_wp_pages %>% 
  mutate(tp_divided_by_page_word_count = talk_page_wordcount / page_wordcount) %>% 
  ggplot(aes(talk_page_wordcount)) +
  geom_histogram() +
  scale_x_log10(labels = scales::comma_format(accuracy = 0.01))
```

Can we detect -ve sentiment on the talk pages? Perhaps as distributions, over every 100 words per page

```{r}
library(tidytext)

What's the typical distribution of the proportion of edits that are reverts?

```{r}
ggplot(ten_k_random_wp_pages,
       aes(rh_revert_prop)) +
  geom_histogram() +
  scale_x_log10()

What are the keywords for controversy in edit messages? And how common are these warrning messages?

```{r}
randoms_revision_history_page_details_ip_warring_words <- 
randoms_revision_history_page_details_ip %>% 
  mutate(warring_words = (str_detect(rh_comment, 'undid|vandalism|why|inaccurate|wrong|warning|blocked|warring|disputed|neutral|unacceptable' )))

# distribution of edits with warring words
randoms_revision_history_page_details_ip_warring_words_prop <- 
randoms_revision_history_page_details_ip_warring_words %>% 
  group_by(page) %>% 
  summarise(prop_warring_edits = sum(warring_words) / n())

ggplot(randoms_revision_history_page_details_ip_warring_words_prop,
         aes(prop_warring_edits)) +
  geom_histogram() +
  scale_x_log10()
```

Is there are correlation between the proportion of warring words and talk page length?

```{r}
# some interesting groupings
randoms_revision_history_page_details_ip_warring_words_prop %>% 
  left_join(ten_k_random_wp_pages %>% 
  mutate(page = 1:nrow(.))) %>% 
  filter(prop_warring_edits > 0) %>% 
  ggplot(aes(prop_warring_edits, 
             talk_page_wordcount ,
             size = page_wordcount)) +
  geom_point(alpha = 0.2)
```

Is there are correlation between the proportion of warring words and revert frequency?

```{r}
randoms_revision_history_page_details_ip_warring_words_prop_revert_prop <- 
randoms_revision_history_page_details_ip_warring_words_prop %>% 
  left_join(ten_k_random_wp_pages %>% 
  mutate(page = 1:nrow(.))) %>% 
  filter(prop_warring_edits != 0,
         rh_revert_prop != 0)

lm(rh_revert_prop ~ prop_warring_edits, 
   data = randoms_revision_history_page_details_ip_warring_words_prop_revert_prop) %>% summary

ggplot(randoms_revision_history_page_details_ip_warring_words_prop_revert_prop,
         aes(prop_warring_edits, 
             rh_revert_prop)) +
  geom_point()
```

time between edits

```{r}
library(lubridate)
randoms_revision_history_page_details_ip_time_diffs <- 
randoms_revision_history_page_details_ip %>% 
  group_nest(page) %>% 
  mutate(time_diff = map(data, ~lag(.x$rh_date) - .x$rh_date)) %>% 
  unnest(time_diff) %>% 
  mutate(time_diff_days = as.numeric(time_diff / 60 / 60 / 24)) %>% 
  mutate(time_diff_hours = as.numeric(time_diff / 60 / 60 )) %>% 
  mutate(less_than_one_month = ifelse(time_diff_hours < 744, TRUE, FALSE)) %>% 
  mutate(less_than_one_week = ifelse(time_diff_hours < 168, TRUE, FALSE)) %>% 
  bind_cols(randoms_revision_history_page_details_ip)

rle_tbl <- 
tibble(lths = rle(randoms_revision_history_page_details_ip_time_diffs$rh_user)$lengths,
vals = rle(randoms_revision_history_page_details_ip_time_diffs$rh_user)$values)

rle_tbl2 <- tibble(
  rh_user = rep(rle_tbl$vals, rle_tbl$lths),
  lths2 = rep(rle_tbl$lths, rle_tbl$lths)
)

randoms_revision_history_page_details_ip_time_diffs %>% 
  filter(!is.na(time_diff_hours)) %>% 
  mutate(quick_three_edits = time_diff_hours + lag(time_diff_hours) + lag(time_diff_hours, 2) + lag(time_diff_hours, 3))

randoms_revision_history_page_details_ip_time_diffs_quick <- 
randoms_revision_history_page_details_ip_time_diffs %>% 
 nest(-page) %>%  
 mutate(edits_3 = map(data, ~.x %>% 
                              filter(less_than_one_week) %>% 
                              filter(n() > 4 ))) %>% 
  unnest(edits_3)

randoms_revision_history_page_details_ip_time_diffs %>% 
  bind_cols(rle_tbl2) %>% 
  filter(lths2 == 1) %>% 
  filter(time_diff_hours < 24)

ggplot(randoms_revision_history_page_details_ip_time_diffs,
       aes(time_diff_days)) +
  geom_histogram()

time series of edits: breakouts relating to events? monthly seems like a good level of aggregation. What can we do for spike detection?

```{r}
randoms_revision_history_page_details_ip_location %>% 
  group_by(page) %>% 
  mutate(the_d = floor_date(rh_date, "day")) %>% 
    mutate(the_m = floor_date(rh_date, "month")) %>% 
  group_by(page, the_m) %>% 
  tally() %>% 
  filter(n() > 130) %>%   # explore this value
  ggplot() +
  geom_line(aes(the_m,
             n)) +
  facet_wrap( ~ page, scales = "free")
```

Spike detection

for randoms, what even is an anomaly?

```{r}
# devtools::install_github("hrbrmstr/AnomalyDetection")
library(AnomalyDetection)
randoms_revision_history_page_details_ip_location_monthly <- 
randoms_revision_history_page_details_ip_location %>% 
  group_by(page) %>% 
  mutate(the_d = floor_date(rh_date, "day")) %>% 
    mutate(the_m = floor_date(rh_date, "month")) %>% 
  group_by(page, the_m) %>% 
  tally() %>% 
  filter(n() > 150) # explore this value

randoms_revision_history_page_details_ip_location_monthly_anom <- 
  randoms_revision_history_page_details_ip_location_monthly %>% 
  ungroup %>% 
  group_nest(page) %>% 
  mutate(anomalies = map(data, ~ad_ts(.x, 
                                      max_anoms = 0.2, 
                                      alpha = 0.005,
                                      threshold = "p95"))) %>% 
  unnest(anomalies)

ggplot() +
  geom_line(
    data=randoms_revision_history_page_details_ip_location_monthly %>% 
      filter(page %in% randoms_revision_history_page_details_ip_location_monthly_anom$page), 
    aes(the_m, n), 
    size=0.125, color="lightslategray")  +
  geom_point(
    data=randoms_revision_history_page_details_ip_location_monthly_anom, 
    aes(timestamp, anoms), 
    color="#cb181d", alpha=1/3) +
  scale_x_datetime(date_labels="%b\n%Y") +
  facet_wrap( ~ page, scales = "free_y")

for WH sites

```{r}
revision_history_page_details_ip_addresses_df_monthly_edits <- 
revision_history_page_details_ip_addresses_df  %>% 
  mutate(the_d = floor_date(rh_date, "day")) %>% 
    mutate(the_m = floor_date(rh_date, "month")) %>% 
  group_by(Site, the_m) %>% 
  tally() %>% 
  filter(n() > 24) # explore this value

# distribution of edits per month
ggplot(revision_history_page_details_ip_addresses_df_monthly_edits,
       aes(n)) +
  geom_histogram() +
  scale_x_log10()

# subset so we get only sites with > 10 edits/month
revision_history_page_details_ip_addresses_df_monthly_edits_10 <- 
revision_history_page_details_ip_addresses_df_monthly_edits %>% 
  filter(max(n) >= 10)

# detect anomalies
revision_history_page_details_ip_addresses_df_monthly_edits_anoms <- 
  revision_history_page_details_ip_addresses_df_monthly_edits_10 %>% 
  ungroup %>% 
  group_nest(Site) %>% 
  mutate(anomalies = map(data, ~ad_ts(.x, 
                                      max_anoms = 0.2, 
                                      alpha = 0.005,
                                      threshold = "p95"))) %>% 
  unnest(anomalies) %>% 
  group_by(Site) %>% 
  filter(max(anoms) >= 50) # where the anomaly is 50 edits/month or more

ggplot() +
  geom_line(
    data=revision_history_page_details_ip_addresses_df_monthly_edits_10 %>% 
      filter(Site %in% revision_history_page_details_ip_addresses_df_monthly_edits_anoms$Site), 
    aes(the_m, n), 
    size=0.125, color="lightslategray")  +
  geom_point(
    data=revision_history_page_details_ip_addresses_df_monthly_edits_anoms, 
    aes(timestamp, anoms), 
    color="#cb181d", alpha=1/3) +
  scale_x_datetime(date_labels="%b\n%Y") +
  facet_wrap( ~ Site, scales = "free")
```

>>>>

get_ip_addresses <- function(x){
  re <- regexpr(
  "(?(?=.*?(\\d+\\.\\d+\\.\\d+\\.\\d+).*?)(\\1|))", 
  x, perl = TRUE)
regmatches(x, re)
}

@@

revision_history_page_details_ip_addresses <- 
revision_history_page_details %>% 
  mutate(edits_from_ip_addresses = map(revision_history_page_details,
                          ~.x %>% 
                            filter(get_ip_addresses(rh_user) != "") %>% 
                            mutate(ip_location = map(rh_user, 
                                                     ~maxmind(.x, ip_db)))  %>% 
                            unnest(ip_location)))

# analyse the countries of IP address vs country of WH site
revision_history_page_details_ip_addresses_df <- 
revision_history_page_details_ip_addresses %>% 
  unnest(edits_from_ip_addresses)

@@

highlight.sector(south_america, track.index = 1, col = "brown", 
    text = "South America", cex = 0.8, text.col = "white", niceFacing = TRUE)
```

@@

time between edits
edit reverts
talk page: length and sentiment
